---
title: "p8105_hw2_mss2284"
author: "Maya Spaur"
date: "9/25/2019"
output: github_document
---
```{r setup, results='hide', warning=FALSE}
library(tidyverse)
library(dplyr)
library(readxl)
```

##Problem 1

Trash Wheel Data:

```{r problem 1 part 1}
trash_wheel_data = 
  read_excel("./Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = "Mr. Trash Wheel") %>%
  janitor::clean_names() %>%
  drop_na(dumpster) %>%
  mutate(
     sports_balls = round(sports_balls),
     sports_balls = as.integer(sports_balls)
  ) %>%
  select(-x15)

```

2017 and 2018 Precipitation Data:

```{r problem 1 part 2}
precipitation_data_2017=
  read_excel("./HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = "2017 Precipitation", skip = 1, col_names = TRUE) %>%
  janitor::clean_names() %>%
  drop_na(total, month) %>%
  mutate(
  year = 2017
)

  
precipitation_data_2018=
  read_excel("./HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = "2018 Precipitation",  skip = 1, col_names = TRUE) %>%
  janitor::clean_names() %>%
  drop_na(total, month) %>%
  mutate(
  year = 2018
)
  

precipitation_data = 
  full_join(precipitation_data_2017, precipitation_data_2018, by = c("month", "total", "year")) %>%
  mutate(
    month = month.name[month]
  )

```

 
The number of observations in Mr. Trash Wheel are `r nrow(trash_wheel_data)`. Examples of key variables include month (spring and summer months), and weight_tons of trash. The median number of sports balls in a dumpster in 2017 was `r median(pull(filter(trash_wheel_data, year==2017), sports_balls))`. 

The number of observations in the Precipitation Data 2017 - 2018 are `r nrow(precipitation_data)`. Key variables are month and total precipitation (total). Among available data, the total precipitation in 2018 was `r sum(pull(precipitation_data_2018, total))`.


##Problem 2

Pols Month Data:

```{r problem 2 part 1, pols-month}
pols_month =
  read_csv("./fivethirtyeight_datasets/pols-month.csv") %>%
  janitor::clean_names() %>%
  separate(mon, into = c("year", "month", "day")) %>%
   mutate(
   year = as.integer(year),
   month = as.integer(month),
   day = as.integer(day),
   month = month.name[month]
   )


pols_month =
  mutate(pols_month,
    president = recode(prez_dem, "0" = "gop", "1" = "dem")
  ) %>%
  select(-day, -prez_dem, -prez_gop)
  
```


Snp Data:

```{r problem 2 part 2, snp}
snp =
  read_csv("./fivethirtyeight_datasets/snp.csv") %>%
  janitor::clean_names() %>%
  separate(date, into = c("month", "day", "year")) %>%
   mutate(
   year = as.integer(year),
   month = as.integer(month),
   day = as.integer(day),
   month = month.name[month]
   ) %>%
  select(year, month, -day, close)
  
```

Unemployment Data:

```{r problem 2 part 3, unemployment}
unemployment =
  read_csv("./fivethirtyeight_datasets/unemployment.csv") %>%
  janitor::clean_names() %>%
  pivot_longer(
    jan:dec,
    names_to = "month",
    values_to = "unemployment_perc"
  ) %>% 
  mutate(month = recode(month, 
    "jan" = "January", 
    "feb" = "February",
    "mar" = "March",
    "apr" = "April",
    "jun" = "June",
    "jul" = "July",
    "aug" = "August",
    "sep" = "September",
    "oct" = "October",
    "nov" = "November",
    "dec" = "December"
   ))

```

Joining the datasets:

```{r problem 2 part 4, joining}
polsmonth_snp = 
  left_join(pols_month, snp, by = c("year", "month")) 

polsmonth_snp_unemployment =
  left_join(polsmonth_snp, unemployment, by = c("year", "month"))
  
```

The pols-month data has `r nrow(pols_month)` observations describing national Democratic or Republican politicians, with a range of years `r range(pull(pols_month, year))`. Key variables include the year and month of the count, and the number of Democratic or Republican presidents (president, created from prez_dem and prez_gop).  

The snp data has `r nrow(snp)` observations describing Standard & Poor's stock market index, including the years `r range(pull(snp, year))`. Key variables include the year and month, and the closing S&P stock value (close) on that date.

The unemployment data has `r nrow(unemployment)` observations describing monthly percentage of unemployment, with a range of years `r range(pull(unemployment, year))`. Key variables include the year and month, and percentage of unemployment (unemployment_perc) for the listed month and year.

The resulting dataset merged the pols_month, snp, and unemployment datasets by year and month, and has `r nrow(polsmonth_snp_unemployment)` observations, with a range of years `r range(pull(polsmonth_snp_unemployment, year))`. Key variables are from the datasets described above, close, unemployment, and president, describing the date of the S&P value at closing, unemployment percentage, and party of the president, respectively.

#Problem 3

Loading and Tidying the Popular Baby Names Data:

```{r problem 3 part 1}
popular_baby_names =
  read_csv("./Popular_Baby_Names.csv") %>%
  janitor::clean_names() %>%
  distinct() %>%
  mutate(
    gender = str_to_lower(gender),
    ethnicity = str_to_lower(ethnicity),
    childs_first_name = str_to_lower(childs_first_name)
  ) %>%
  mutate(
    ethnicity = replace(ethnicity, ethnicity == "black non hisp" , "black non hispanic"),
    ethnicity = replace(ethnicity, ethnicity == "white non hisp" , "white non hispanic"),
    ethnicity = replace(ethnicity, ethnicity == "asian and paci" , "asian and pacific islander")
    )

```

Olivia Popularity:

```{r problem 3 part 2}

olivia_data = 
  filter(popular_baby_names, childs_first_name == "olivia") %>%
  select(-gender, -count, -childs_first_name) %>%
  pivot_wider(
  names_from = "year_of_birth",
  values_from = "rank"
)

olivia_data

```

Top 2016 Names for White Non Hispanic Males:

```{r problem 3 part 3}
male_names = 
  filter(popular_baby_names, gender == "male", rank=="1")%>%
  select(-count, -gender, -rank) %>%
  pivot_wider(
    names_from = "year_of_birth",
  values_from = "childs_first_name"
  )
  
white_nonhisp_2016=
  filter(popular_baby_names, gender == "male", ethnicity == "white non hispanic", year_of_birth == "2016") %>%
  ggplot(aes(x = rank, y = count)) + geom_point()
  
white_nonhisp_2016
    
```

